---
title: "LRcellTypeMarkers: Marker gene information for LRcell R Bioconductor package using Bioconductor's ExperimentHub"
author: "Wenjing Ma"
date: "`r doc_date()`"
bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{LRcellTypeMarkers: Marker gene information for LRcell R Bioconductor package using Bioconductor's ExperimentHub}
  %\VignetteEngine{knitr::rmarkdown}
output: 
  BiocStyle::html_document
---

# Download cell-type specific marker gene information

LRcellTypeMarkers currently provides gene enrichment scores for mouse whole brain
and human prefrontal cortex. These gene enrichment scores were calculated
using the algorithm proposed in [@marques2016oligodendrocyte]. The detailed 
procedures are described in `scripts/make-data.R`. Users are encouraged
to download the data, generate customized cell-type specific marker genes and
apply it in on bulk DE genes using [LRcell](https://github.com/marvinquiet/LRcell). 

The example below shows how to download the data using ExperimentHub.

```{r intro}
library(ExperimentHub)
eh = ExperimentHub()
query(eh, "LRcellTypeMarkers")
```

According to the description, users can choose the most suitable marker gene
information for further analysis. Here, we used mouse frontal cortex marker genes
as an example, which has the title as **EH4548**.
```{r extract_data}
mouse_FC <- eh[['EH4548']]
mouse_FC[1:6, 1:6]  # show head of the data
```

From the data, we can observe that the rows are gene symbols and the columns are 
the sub-cell types (or clusters). The values indicate the gene enrichment level 
in a sub-cell type (or cluster). Higher the values are, more unique the genes 
are in the certain cell types. Users can sort the enrichment score in a 
descending order and select top 100 genes for each sub-cell type (or cluster) 
using the following command or the function named `get_markergenes()` in LRcell.

Here, we take a glance at the first 6 clusters and the first 6 marker genes in 
each cluster.

```{r sort}
FC_marker_genes <- list()

for (cluster in colnames(mouse_FC)) {
  enriched_values <- as.numeric(mouse_FC[, cluster])
  names(enriched_values) <- rownames(mouse_FC)
  sorted_enriched_genes <- sort(enriched_values, decreasing = TRUE)
  FC_marker_genes[[cluster]] <- head(names(sorted_enriched_genes), n=100)
}

head(lapply(FC_marker_genes, head))  # glance at the marker genes
```

For further usage in LRcell, please check [LRcell Tutorial](https://github.com/marvinquiet/LRcell).

# sessionInfo()

```{r sessionInfo}
sessionInfo()
```
